Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-12 paper/paper.Rmd
done with spatial displays

@@

<!-- Many sets of Australian statistical areas have a large disparity between the smallest and largest units.
The rural communities in Australia operate on a much larger geographic space than small inner-city communities. When a spatial distribution is related to the size of the areas, or the population density, the size of the regions can allow erroneous conclusions to be drawn about the state of the statistic over the entire population, when presenting the statistics on a geographic map base. This occurs as large regions filled with a consistent colour or pattern can draw the attention of map readers, and small regions are not paid equal attention [@CTTMB]. An alternative display can be used to honestly communicate a spatial distribution for a set of heterogeneous areas [@NISCC]. 
>>>>>>> de6ff98be8766902375ed1f9fcdcea82b768d1ea
This work aims to show that a hexagon tile map display is a viable alternative to the geographic map base for presenting population statistics.

@@

<!-- Tesselated tile gram -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by Raisz [-@RSCW], using rectangles that tile to align borders of neighbours [@CDWCS]. A thorough review of the array of methods, as suitable for cancer atlas displays is available in XXX.

```{r liver, fig.height = 9, fig.width = 6, fig.cap = "The smoothed average of liver cancer diagnoses for Australian males. The divering colour scheme shows dark blue areas with much lower than average diagnoses, yellow areas with diagnoses around the Australian average, orange with above average and red shows diagnoses much higher than average.", results = "asis"}
ggdraw(ylim = c(0,2), xlim = c(0,1.3), clip = "on") +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m.png")), -0.15, 0.9, 1.6, 1.2) +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m_hex.png")), -0.15, -0.1, 1.6, 1.2) +
  draw_plot_label(c("a", "b"), 
    c(0.05, 0.05), c(1.95, 1), size = 20, color = "white") + 
  theme(plot.background = element_rect(fill = "black"))
```
<!--
This will create the figure as needed in the paper, but I can't get the width argument to work properly.
\begin{figure*}[!t]
\centering
\includegraphics{paper_files/figure-latex/liver-1.pdf}
\caption{A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.}
\label{fig:liver2}
\end{figure*}
-->

<!-- Specific problem for Australia -->
Australia is an extreme case of heterogeneous geographic units, with the large differences between the smallest and largest geographic units across various granularities.
To communicate information regarding the small inner-city geographic units of Australian cities, an alternative display must be employed. This will emphasize the value of the statistic for areas with a high population. Fig. \ref{fig:liver} shows the distribution of the smoothed average amount of liver cancer diagnoses for Australian Males from 2005 to 2014.
These inner-city areas are not visible on the choropleth map (a). The hexagon tile map (b) shows orange hexagons for the inner-city SA2s with higher than average levels of diagnoses in the capital cities of Melbourne and Sydney. The northern Queensland and Northern Territory SA2s with higher than average rates are still visible but use less map space on the hexagon tile map. There are still many light blue areas in both displays.

>>>>

<!-- Many sets of Australian statistical areas have a large disparity between the smallest and largest units.
The rural communities in Australia operate on a much larger geographic space than small inner-city communities. When a spatial distribution is related to the size of the areas, or the population density, the size of the regions can allow erroneous conclusions to be drawn about the state of the statistic over the entire population, when presenting the statistics on a geographic map base. This occurs as large regions filled with a consistent colour or pattern can draw the attention of map readers, and small regions are not paid equal attention [@CTTMB]. An alternative display can be used to honestly communicate a spatial distribution for a set of heterogeneous areas [@NISCC]. 
This work aims to show that a hexagon tile map display is a viable alternative to the geographic map base for presenting population statistics.

```{r liver, fig.height = 6, fig.width = 4, fig.cap = "A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.", results = "asis"}
ggdraw(ylim = c(0,2), xlim = c(0,1.3), clip = "on") +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m.png")), -0.15, 0.9, 1.6, 1.2) +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m_hex.png")), -0.15, -0.1, 1.6, 1.2) +
  draw_plot_label(c("a", "b"), 
    c(0.05, 0.05), c(1.95, 1), size = 20, color = "white") + 
  theme(plot.background = element_rect(fill = "black"))
```
<!--
This will create the figure as needed in the paper, but I can't get the width argument to work properly.
\begin{figure*}[!t]
\centering
\includegraphics{paper_files/figure-latex/liver-1.pdf}
\caption{A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.}
\label{fig:liver2}
\end{figure*}
-->

@@

@@

<!-- Other cartograms -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by Raisz [-@RSCW], using rectangles that tile to align borders of neighbours [@CDWCS]. A thorough review of the array of methods, as suitable for cancer atlas displays is available in XXX.

<!-- Hexagon tile map -->
The hexagon tile map algorithm, autmatically matches spatial regions to their nearest hexagon tile, from a grid of tiles. It has the effect of spreading out the inner city areas while maintaining the spatial locations or regions in remote areas. The algorithm is available in the R package, sugarbag (XXX REF). Figure \ref{fig:liver} shows the hexagon tile map, along with the choropleth map of liver cancer rates in Australia. Colour maps from substantially below average (blue) to substantially above average (red) rates. The inner city areas are expanded out, making it possible to see the cancer incidence in the small, densely populated areas. Remote regions are represented by isolated hexagons, which is not ideal, but maintains the spatial location of these data values. It is oof interest to know how well the spatial distribution is perceived for this display, in comparison to the choropleth.

<!-- Specific problem for Australia 
Australia is an extreme case of heterogeneous geographic units, with the large differences between the smallest and largest geographic units across various granularities.
To communicate information regarding the small inner-city geographic units of Australian cities, an alternative display must be employed. This will emphasize the value of the statistic for areas with a high population. Fig. \ref{fig:liver} shows the distribution of the smoothed average amount of liver cancer diagnoses for Australian Males from 2005 to 2014.
These inner-city areas are not visible on the choropleth map (a). The hexagon tile map (b) shows orange hexagons for the inner-city SA2s with higher than average levels of diagnoses in the capital cities of Melbourne and Sydney. The northern Queensland and Northern Territory SA2s with higher than average rates are still visible but use less map space on the hexagon tile map. There are still many light blue areas in both displays. 
-->

### 2020-01-12 paper/paper.tex
done with spatial displays

Australia is an extreme case of heterogeneous geographic units, with the large differences between the smallest and largest geographic units across various granularities.
To communicate information regarding the small inner-city geographic units of Australian cities, an alternative display must be employed. This will emphasize the value of the statistic for areas with a high population. Fig. \ref{fig:liver} shows the distribution of the smoothed average amount of liver cancer diagnoses for Australian Males from 2005 to 2014.
These inner-city areas are not visible on the choropleth map (a). The hexagon tile map (b) shows orange hexagons for the inner-city SA2s with higher than average levels of diagnoses in the capital cities of Melbourne and Sydney. The northern Queensland and Northern Territory SA2s with higher than average rates are still visible but use less map space on the hexagon tile map. There are still many light blue areas in both displays.

>>>>

The hexagon tile map algorithm, autmatically matches spatial regions to their nearest hexagon tile, from a grid of tiles. It has the effect of spreading out the inner city areas while maintaining the spatial locations or regions in remote areas. The algorithm is available in the R package, sugarbag (XXX REF). Figure \ref{fig:liver} shows the hexagon tile map, along with the choropleth map of liver cancer rates in Australia. Colour maps from substantially below average (blue) to substantially above average (red) rates. The inner city areas are expanded out, making it possible to see the cancer incidence in the small, densely populated areas. Remote regions are represented by isolated hexagons, which is not ideal, but maintains the spatial location of these data values. It is oof interest to know how well the spatial distribution is perceived for this display, in comparison to the choropleth.
